---
title: "TSA"
author: "Yuxiang Ren"
output: pdf_document
always_allow_html: true
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: sentence
---
#packges    
```{r packages}
library(readxl)
#library(xlsx) #this package doesn't work on Macs very well
library(tidyr)
library(lubridate)
library(ggplot2)
library(forecast)  
library(Kendall)
library(tseries)
library(outliers)

library(tidyverse)
library(smooth)
library(zoo)
library(kableExtra)
library(dplyr)
library(smooth)

```
#data
```{r}
raw <- read.csv("./Data/Processed/PM2.5_daily_city_2000_2021.csv")
Table_raw <- raw[c(1:20),]
kable(Table_raw,caption = "Raw data")

raw_wider <- raw %>% pivot_wider(names_from = city_id, values_from = meanpm) %>% 
  mutate(date = as.Date(paste(year, month, day, sep = "-"))) %>%
  select(date, everything()) 

```
#Beijing 1100
##data
```{r}
Beijing <- raw_wider[,c('date', 'year', 'month', 'day', '1100')] %>% rename(pm2.5 = '1100') %>% group_by(year, month) %>% summarize(monthlyPM = mean(pm2.5)) %>% mutate(date = as.Date(paste(year, month, 01, sep = "-")))

#question, mean PM or accumulated PM

Beijing_train <- Beijing %>% filter(date < as.Date("2021-01-01"))
Beijing_test <- Beijing %>% filter(date >= as.Date("2021-01-01"))

ts_Beijing_train <- ts(Beijing_train$monthlyPM,
                           start=c(2000,1), frequency = 12)
ts_Beijing_all <- ts(Beijing$monthlyPM,
                           start=c(2000,1), frequency = 12)
```
##Model 1: Arithmetic mean
```{r}
AM_Beijing <- meanf(y = ts_Beijing_train, h = 12)  
checkresiduals(AM_Beijing)
plot(AM_Beijing)
```
## Model 2: Seasonal naive
```{r}
SNAIVE_Beijing <- snaive(ts_Beijing_train, h=12, holdout=FALSE)
checkresiduals(SNAIVE_Beijing)
plot(SNAIVE_Beijing)
```
## Model 3: SARIMA
```{r}
SARIMA_Beijing_fit <- auto.arima(ts_Beijing_train)
checkresiduals(SARIMA_Beijing_fit)

#Generating forecasts
#remember auto.arima does not call the forecast() internally so we need one more step
SARIMA_Beijing <- forecast(SARIMA_Beijing_fit,h=12)
plot(SARIMA_Beijing)
```
## Model 4: seasonal exponential smoothing，Triple Exponential Smoothing
```{r}
SSES_Beijing <- es(ts_Beijing_train,model="ZZZ",h=12,holdout=FALSE)
plot(SSES_Beijing)
checkresiduals(SSES_Beijing)

```
##Model 2+3
```{r}
SNAIVE_Beijing$mean
SARIMA_Beijing$mean

df23 <- data.frame(SNAIVE_Beijing$mean, SARIMA_Beijing$mean)
colnames(df23) <- c("SNAIVE","SARIMA")

df23_mean <- df23 %>% mutate(SNAIVE_SARIMA = (SNAIVE + SARIMA)/2)
```

##Model 3+4
```{r}
SSES_Beijing$forecast
SARIMA_Beijing$mean

df24 <- data.frame(SSES_Beijing$forecast, SARIMA_Beijing$mean)
colnames(df24) <- c("SSES","SARIMA")

df24_mean <- df24 %>% mutate(SSES_SARIMA = (SSES + SARIMA)/2)
```


##accuacy  
```{r accuacy}
MEAN_scores <- accuracy(AM_Beijing$mean,Beijing_test$monthlyPM)
SNAIVE_scores <- accuracy(SNAIVE_Beijing$mean,Beijing_test$monthlyPM)
SARIMA_scores <- accuracy(SARIMA_Beijing$mean,Beijing_test$monthlyPM)
SSES_scores <- accuracy(SSES_Beijing$forecast,Beijing_test$monthlyPM)
M2_3_scores <- accuracy(df23_mean$SNAIVE_SARIMA,Beijing_test$monthlyPM)
M2_4_scores <- accuracy(df24_mean$SSES_SARIMA,Beijing_test$monthlyPM)
# 4 models Table
Beijing_scores0 <- as.data.frame(rbind(MEAN_scores, SNAIVE_scores, SARIMA_scores,SSES_scores))

row.names(Beijing_scores0) <- c("MEAN", "SNAIVE","SARIMA","SSES")

kbl(Beijing_scores0, 
      caption = "Forecast Accuracy for Seasonal Data",
      digits = array(4,ncol(Beijing_scores0))) %>%
  kable_styling(full_width = FALSE, position = "center") %>%
  #highlight model with lowest RMSE
  kable_styling(latex_options="striped", stripe_index = which.min(Beijing_scores[,"RMSE"]))
##save
write.csv(Beijing_scores0, row.names = FALSE, 
          file = "./Data/Processed/Beijing_scores0.csv")
# 6 models table
Beijing_scores <- as.data.frame(rbind(MEAN_scores, SNAIVE_scores, SARIMA_scores,SSES_scores, M2_3_scores,M2_4_scores))
row.names(Beijing_scores) <- c("MEAN", "SNAIVE","SARIMA","SSES","SNAIVE_SARIMA","SSES_SARIMA")

kbl(Beijing_scores, 
      caption = "Forecast Accuracy for Seasonal Data",
      digits = array(6,ncol(Beijing_scores))) %>%
  kable_styling(full_width = FALSE, position = "center") %>%
  #highlight model with lowest RMSE
  kable_styling(latex_options="striped", stripe_index = which.min(Beijing_scores[,"RMSE"]))

write.csv(Beijing_scores, row.names = FALSE, 
          file = "./Data/Processed/Beijing_scores.csv")

best_model_index <- which.min(Beijing_scores[,"RMSE"])
cat("The best model by RMSE is:", row.names(Beijing_scores[best_model_index,])) 

```
##test result-2023
###SSES model, two year result are same
```{r SSES}
for_Beijing_result <- es(ts_Beijing_all,model="ZZZ",h=24,holdout=FALSE)
plot(SSES_Beijing)
forecast_Beijing <- for_Beijing_result$forecast
```
###SNAIVE same two years
```{r SNAIVE}
for_Beijing_result2 <- snaive(ts_Beijing_all, h=48, holdout=FALSE)
plot(for_Beijing_result2)
for_Beijing_result2$mean
```
###SARIMA
can forecast several years 
```{r SARIMA}
for_Beijing_result3_fit <- auto.arima(ts_Beijing_all)
for_Beijing_result3 <- forecast(for_Beijing_result3_fit,h=24)
plot(for_Beijing_result3)
for_Beijing_result3$mean

```
##final
use SSES+SARIMA
###beijing 1100
```{r final_Beijing}
#data
Beijing <- raw_wider[,c('date', 'year', 'month', 'day', '1100')] %>% rename(pm2.5 = '1100') %>% group_by(year, month) %>% summarize(monthlyPM = mean(pm2.5)) %>% mutate(date = as.Date(paste(year, month, 01, sep = "-")))
#ts.
ts_Beijing <- ts(Beijing$monthlyPM,start=c(2000,1), frequency = 12)
#SSES
SSES_Beijing <- es(ts_Beijing,model="ZZZ",h=60,holdout=FALSE)
#SARIMA
SARIMA <- auto.arima(ts_Beijing, seasonal = TRUE)
SARIMA_Beijing <- forecast(SARIMA,h=60)
#result
R_Beijing <- (SSES_Beijing$forecast+SARIMA_Beijing$mean)/2

```

###Shanghai 3100
```{r final_Shanghai}
#data
Shanghai <- raw_wider[,c('date', 'year', 'month', 'day', '3100')] %>% rename(pm2.5 = '3100') %>% group_by(year, month) %>% summarize(monthlyPM = mean(pm2.5)) %>% mutate(date = as.Date(paste(year, month, 01, sep = "-")))
#ts.
ts_Shanghai <- ts(Shanghai$monthlyPM,start=c(2000,1), frequency = 12)
#SSES
SSES_Shanghai <- es(ts_Shanghai,model="ZZZ",h=60,holdout=FALSE)
#SARIMA
SARIMA <- auto.arima(ts_Shanghai, seasonal = TRUE)
SARIMA_Shanghai <- forecast(SARIMA,h=60)
#result
R_Shanghai <- (SSES_Shanghai$forecast+SARIMA_Shanghai$mean)/2

```

###Guangzhou 4401
```{r final_Guangzhou}
#data
Guangzhou <- raw_wider[,c('date', 'year', 'month', 'day', '4401')] %>% rename(pm2.5 = '4401') %>% group_by(year, month) %>% summarize(monthlyPM = mean(pm2.5)) %>% mutate(date = as.Date(paste(year, month, 01, sep = "-")))
#ts.
ts_Guangzhou <- ts(Guangzhou$monthlyPM,start=c(2000,1), frequency = 12)
#SSES
SSES_Guangzhou <- es(ts_Guangzhou,model="ZZZ",h=60,holdout=FALSE)
#SARIMA
SARIMA <- auto.arima(ts_Guangzhou, seasonal = TRUE)
SARIMA_Guangzhou <- forecast(SARIMA,h=60)
#result
R_Guangzhou <- (SSES_Guangzhou$forecast+SARIMA_Guangzhou$mean)/2

```

###Shenzhen 4403
```{r final_Shenzhen}
#data
Shenzhen <- raw_wider[,c('date', 'year', 'month', 'day', '4403')] %>% rename(pm2.5 = '4403') %>% group_by(year, month) %>% summarize(monthlyPM = mean(pm2.5)) %>% mutate(date = as.Date(paste(year, month, 01, sep = "-")))
#ts.
ts_Shenzhen <- ts(Shenzhen$monthlyPM,start=c(2000,1), frequency = 12)
#SSES
SSES_Shenzhen <- es(ts_Shenzhen,model="ZZZ",h=60,holdout=FALSE)
#SARIMA
SARIMA <- auto.arima(ts_Shenzhen, seasonal = TRUE)
SARIMA_Shenzhen <- forecast(SARIMA,h=60)
#result
R_Shenzhen <- (SSES_Shenzhen$forecast+SARIMA_Shenzhen$mean)/2

```

###Tianjin 1200
```{r final_Tianjin}
#data
Tianjin <- raw_wider[,c('date', 'year', 'month', 'day', '1200')] %>% rename(pm2.5 = '1200') %>% group_by(year, month) %>% summarize(monthlyPM = mean(pm2.5)) %>% mutate(date = as.Date(paste(year, month, 01, sep = "-")))
#ts.
ts_Tianjin <- ts(Tianjin$monthlyPM,start=c(2000,1), frequency = 12)
#SSES
SSES_Tianjin <- es(ts_Tianjin,model="ZZZ",h=60,holdout=FALSE)
#SARIMA
SARIMA <- auto.arima(ts_Tianjin, seasonal = TRUE)
SARIMA_Tianjin <- forecast(SARIMA,h=60)
#result
R_Tianjin <- (SSES_Tianjin$forecast+SARIMA_Tianjin$mean)/2

```

###Chongqing 1100
```{r final_Chongqing}
#data
Chongqing <- raw_wider[,c('date', 'year', 'month', 'day', '5000')] %>% rename(pm2.5 = '5000') %>% group_by(year, month) %>% summarize(monthlyPM = mean(pm2.5)) %>% mutate(date = as.Date(paste(year, month, 01, sep = "-")))
#ts.
ts_Chongqing <- ts(Chongqing$monthlyPM,start=c(2000,1), frequency = 12)
#SSES
SSES_Chongqing <- es(ts_Chongqing,model="ZZZ",h=60,holdout=FALSE)
#SARIMA
SARIMA <- auto.arima(ts_Chongqing, seasonal = TRUE)
SARIMA_Chongqing <- forecast(SARIMA,h=60)
#result
R_Chongqing <- (SSES_Chongqing$forecast+SARIMA_Chongqing$mean)/2

```

###Chengdu 5101
```{r final_chengdu}
#data
Chengdu <- raw_wider[,c('date', 'year', 'month', 'day', '5101')] %>% rename(pm2.5 = '5101') %>% group_by(year, month) %>% summarize(monthlyPM = mean(pm2.5)) %>% mutate(date = as.Date(paste(year, month, 01, sep = "-")))
#ts.
ts_Chengdu <- ts(Chengdu$monthlyPM,start=c(5101,1), frequency = 12)
#SSES
SSES_Chengdu <- es(ts_Chengdu,model="ZZZ",h=60,holdout=FALSE)
#SARIMA
SARIMA <- auto.arima(ts_Chengdu, seasonal = TRUE)
SARIMA_Chengdu <- forecast(SARIMA,h=60)
#result
R_Chengdu <- (SSES_Chengdu$forecast+SARIMA_Chengdu$mean)/2

```

###all
```{r all_save}
all_forcast <- data.frame(R_Beijing, R_Shanghai, R_Guangzhou, R_Shenzhen, R_Tianjin, R_Chongqing)
date_sequence <- seq(from = as.Date("2022-01-01"), by = "month", length.out = 60)
all_forcast$Date <- date_sequence
all_forcast <- all_forcast[,c(7,1:6)]
write.csv(all_forcast, row.names = FALSE, 
          file = "./Data/Processed/all_forcast.csv")

```
#seasonal+trend
```{r seasonal +trend}
stl_Beijing <- stl(ts_Beijing, s.window = "periodic")
stl_Shanghai <- stl(ts_Shanghai, s.window = "periodic")
stl_Guangzhou <- stl(ts_Guangzhou, s.window = "periodic")
stl_Shenzhen <- stl(ts_Shenzhen, s.window = "periodic")
stl_Tianjin <- stl(ts_Tianjin, s.window = "periodic")
stl_Chongqing <- stl(ts_Chongqing, s.window = "periodic")
stl_Chengdu <- stl(ts_Chengdu, s.window = "periodic")

seasonal_Beijing <- stl_Beijing$time.series[c(1:12), "seasonal"]
seasonal_Shanghai <- stl_Shanghai$time.series[c(1:12), "seasonal"]
seasonal_Guangzhou <- stl_Guangzhou$time.series[c(1:12), "seasonal"]
seasonal_Shenzhen <- stl_Shenzhen$time.series[c(1:12), "seasonal"]
seasonal_Tianjin <- stl_Tianjin$time.series[c(1:12), "seasonal"]
seasonal_Chongqing <- stl_Chongqing$time.series[c(1:12), "seasonal"]
seasonal_Chengdu <- stl_Chengdu$time.series[c(1:12), "seasonal"]

trend_Beijing <- stl_Beijing$time.series[, "trend"]
trend_Shanghai <- stl_Shanghai$time.series[, "trend"]
trend_Guangzhou <- stl_Guangzhou$time.series[, "trend"]
trend_Shenzhen <- stl_Shenzhen$time.series[, "trend"]
trend_Tianjin <- stl_Tianjin$time.series[, "trend"]
trend_Chongqing <- stl_Chongqing$time.series[, "trend"]
trend_Chengdu <- stl_Chengdu$time.series[, "trend"]

```
##seasonalALL
```{r seasonalALL}
seasonal_data <- data.frame(
  city = rep(c("Beijing", "Shanghai", "Guangzhou", "Shenzhen", "Tianjin", "Chongqing", "Chengdu"), each = 12),
  month = rep(month.abb, 7),
  seasonal = c(seasonal_Beijing, seasonal_Shanghai, seasonal_Guangzhou, seasonal_Shenzhen, seasonal_Tianjin, seasonal_Chongqing, seasonal_Chengdu)
)
#plot1_seasonal,"not good
plot1 <- ggplot(seasonal_data, aes(x = month, y = seasonal, group = city)) +
  geom_line() +
  scale_x_discrete(limits = month.abb) +
  facet_wrap(~city, nrow = 3) +
  labs(title = "Seasonal Patterns in 7 Chinese Cities", x = "Month", y = "Seasonal Component") +
  theme_minimal()

#plot2_seasonal,
plot2 <- ggplot(seasonal_data, aes(x = month, y = seasonal, group = city, colour = city)) +
  geom_line() +
  scale_x_discrete(limits = month.abb) +
  labs(title = "Figure 2. Seasonality in PM2.5 in 7 Chinese Cities", x = "Month", y = "PM2.5 Seasonal Component") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  guides(colour = guide_legend(title = "City"))


write.csv(seasonal_data, row.names = FALSE, 
          file = "./Data/Processed/seasonal_data.csv")

```
##trendALL

```{r trend_all}
date_seq <- seq(as.Date("2000-01-01"), as.Date("2021-12-01"), by = "month")
trend_data <- data.frame(
  city = rep(c("Beijing", "Shanghai", "Guangzhou", "Shenzhen", "Tianjin", "Chongqing", "Chengdu"), each = length(date_seq)),
  date = rep(date_seq, 7),
  trend = c(trend_Beijing, trend_Shanghai, trend_Guangzhou, trend_Shenzhen, trend_Tianjin, trend_Chongqing, trend_Chengdu)
)

ggplot(trend_data, aes(x = date, y = trend, group = city, colour = city)) +
  geom_line() +
  labs(title = "Figure 3. Trends in PM2.5 in 7 Chinese Cities", x = "Date", y = "PM2.5 Trend Component") +
  scale_x_date(breaks = "1 year", date_labels = "%Y") +
  theme_minimal() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(colour = guide_legend(title = "City"))

write.csv(trend_data, row.names = FALSE, 
          file = "./Data/Processed/trend_data.csv")


```


